---
title: "Welcome My World"
---

Hello, my name is Xu Tang. I graduated with a Bachelor’s degree in Economics from the Beijing University of Posts and Telecommunications. I have experience in sales, business development, and data analysis, currently working at WowRobee Inc. in New York, where I focus on customer analysis and business strategy for delivery robots.

Previously, I interned at Deloitte and CITIC Securities, gaining research and financial analysis experience. I am fluent in Chinese and English, skilled in Excel, Python, and SQL, and I enjoy tennis and video editing in my free time.


```{r}
#| include: false
1+1
```
--------------
Last Updated: `r format(Sys.time(), "%A %m %d, %Y at %H:%M%p")`

```{r}
#| echo: false
#| message: false
#| warning: false

if(!require("leaflet")){
    options(repos=c(CRAN="https://cloud.r-project.org"))
    install.packages("leaflet")
    stopifnot(require("leaflet"))
}

baruch_longitude <- -73.98333
baruch_latitude  <- +40.74028

leaflet() |>
  addTiles() |>
  setView(baruch_longitude, baruch_latitude, zoom=17) |>
  addPopups(baruch_longitude, baruch_latitude, 
            "I am a Master's student at <b>Baruch College</b>!")
```